---
title: "Inferential Data Analysis"
author: "Alberto Macías"
date: "9/29/2020"
output: html_document
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

We realize and inferential data analysis on the ToothGrowth data set in order to study how vitamin C provided via orange juice or via ascorbic acid affects in the tooth length in 60 guinea pigs.

## Loading the data and performing the EDA


```{r}
library(datasets)
data("ToothGrowth")
str(ToothGrowth)
```

According to R documentation, Toothgrowth data set contains data about the "length of odontoblast (which are cells responsible for tooth growth) in 60 guinea pigs". It is a data frame with 4 variables and 60 observations, one for each guinea pig. The variables are len, a numeric vector that contains the lengths of the cells in unspecified units; supp, a two level factor that indicates the supplement type of vitamin C: orange juice (OJ) or ascorbic acid (VC); and dose, a numeric vector that contains the dosage in which vitamin C was provided, in  3 dosages, 0.5, 1 and 2 milligrams per day (mg/day).

We divide ToothGrowth data set into to data frames, vc and oj, divided by the supplement type. We compute mean, standard deviation and variance of the cell lengths

```{r}
vc <- subset(ToothGrowth, supp=="VC")
oj <- subset(ToothGrowth, supp=="OJ")
mnvc <- mean(vc$len)
sdvc <- sd(vc$len)
varvc <- var(vc$len)
mnoj <- mean(oj$len)
sdoj <- sd(oj$len)
varoj <- var(oj$len)
```





