---
title: "Inferential Data Analysis"
author: "Alberto Macías"
date: "9/29/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

We realize and inferential data analysis on the ToothGrowth data set in order to study how vitamin C provided via orange juice or via ascorbic acid affects in the tooth length in 60 guinea pigs.

## Loading the data and performing the EDA

```{r, results='hide', message=FALSE}
library(ggplot2)
library(dplyr)
```


```{r}
library(datasets)
data("ToothGrowth")
str(ToothGrowth)
```

According to R documentation, Toothgrowth data set contains data about the "length of odontoblast (which are cells responsible for tooth growth) in 60 guinea pigs". It is a data frame with 3 variables and 60 observations, one for each guinea pig. The variables are len, a numeric vector that contains the lengths of the cells in unspecified units; supp, a two level factor that indicates the supplement type of vitamin C: orange juice (OJ) or ascorbic acid (VC); and dose, a numeric vector that contains the dosage in which vitamin C was provided, in  3 dosages, 0.5, 1 and 2 milligrams per day (mg/day).

We divide ToothGrowth data set into to data frames, vc and oj, divided by the supplement type. We compute mean, standard deviation and variance of the cell lengths

```{r}
vc <- subset(ToothGrowth, supp=="VC")
oj <- subset(ToothGrowth, supp=="OJ")
mnvc <- mean(vc$len); varvc <- var(vc$len)
mnoj <- mean(oj$len); varoj <- var(oj$len)
mnvc; varvc;
mnoj; varoj

```
We note that the mean of length of cells in oj data is greater than mean of length in vc, which could indicate that the orange juice supplement of vitamin c affects more to cell length than ascorbic acid. Also that data in the oranjuice set is less variable.





## Appendix



```{r, fig.width=3, fig.height=2}

ggplot(ToothGrowth, aes(x=supp, y=len, fill= supp))+
            geom_boxplot()+
            ggtitle("Length of Odontoblast by supplement")+
            ylab("Length")
            
ggplot(vc, aes(x=dose, y=len, fill=dose, group = dose))+
            geom_boxplot()+
            ggtitle("Length of Odontoblast by dose of ascorbic acid.")+
            ylab("Length")

ggplot(oj, aes(x=dose, y=len, fill=dose, group = dose))+
            geom_boxplot()+
            ggtitle("Length of Odontoblast by dose of orange juice.")+
            ylab("Length")

```